---
title: "Stat 632 Final Project"
author: 'Michael Pantoja Collasso, Yuanyuan Fan, Wennie Wang'
date: "2023-04-27"
output: pdf_document
---

### Interested in Predict 5-Year Career Longevity for NBA Rookies y = 0 if career years played < 5 y = 1 if career years played >= 5

resource: data.world
<https://data.world/exercises/logistic-regression-exercise-1>

### Description:

Name: names for NBA players 

GP: games played 

MIN: minutes played 

PTS:points per game

FGM: field goals made 

FGA: field goal attempts 

FG.: field goal percent 

X3P.Made: 3 points made 

X3PA: 3 points attempts 

X3P.: 3 points percent 

FTM: free throw amde 

FTA: free throw attempts 

FT.: free throw percent 

OREB: offensice rebounds 

DREB: defensice rebounds 

AST: assists 

STL: steals 

BLK: blocks 

TOV: turnovers 

TARGET_5Yrs: outcome, 1 if career length >=5 yrs, 0 if career length < 5 yrs

#### Full Model: 
$$log[\frac{P({Target5Yrs})} {1-P({Target5Yrs})}] =\beta_{0}+\beta_{1}GP+\beta_{2}MIN+...+\beta_{18}BLK+\beta_{19}TOV+ \epsilon$$
#### Hypothese: 

$H_0: \beta_1 = \beta_2 = ...= \beta_{19} = 0$

$H_1: \beta_i\neq0, (i = 1,2,3,...19)$

significant level at $\alpha = 0.05$

#### Estimated Model
$log[\frac{P({Target5Yrs})} {1-P({Target5Yrs})}]=-4.22+0.04GP-0.05MIN+0.09FGA+0.02FG.+$

$3.15X3P.Made-1.06X3PA++0.02FT.-1.02DREB+0.86REB+0.25AST +0.56BLK$


```{r}
library(car)
library(ggplot2)
library(pROC)
NBA <- read.csv("nba.csv")
```

```{r}
NBA$TARGET_5Yrs <- as.factor(NBA$TARGET_5Yrs)
```

```{r}
NBA <- na.omit(NBA)
summary(NBA)
```

```{r message = FALSE}
# scatter Matrix
library(GGally)

# Subset the data to predictor variables and the response variable
predictors1 <- NBA[, 2:6]
response <- NBA$TARGET_5Yrs

# Combine predictors and response variable into one data frame
data1 <- cbind(predictors1, response)

# Create the scatterplot matrix with GGally
ggpairs(data = data1, columns = 1:5, aes(color = response))
```

```{r message = FALSE}
predictors2 <- NBA[, 7:12]
data2 <- cbind(predictors2, response)
ggpairs(data = data2, columns = 1:6, aes(color = response))
```


```{r message = FALSE}
predictors3 <- NBA[, 13:16]
data3 <- cbind(predictors3, response)
ggpairs(data = data3, columns = 1:4, aes(color = response))
```


```{r message = FALSE}
predictors4 <- NBA[, 17:20]
data4 <- cbind(predictors4, response)
ggpairs(data = data4, columns = 1:4, aes(color = response))
```

```{r}
model1 <- glm(TARGET_5Yrs ~ . - Name, data = NBA, family = "binomial")
summary(model1)
```

```{r}
model2 <- stats::step(model1, trace = 0)
```

```{r}
summary(model2)
```

```{r}
set.seed(123)
```

```{r}
n <- nrow(NBA); n
```

```{r}
floor(0.7*n)
```

```{r}
train <- sample(1:n, 930)
```

```{r}
glm_train <- glm(
                 TARGET_5Yrs ~ GP + MIN + FGA + FG. + X3P.Made + X3PA + FT. + DREB + REB + AST + BLK, 
                 data=NBA, subset = train, family = binomial)
```

```{r}
summary(glm_train)
```

```{r}
NBA_test <- NBA[-train, ]
```

```{r}
probs_test <- predict(glm_train, newdata = NBA_test, type = "response")
```

```{r}
length(probs_test)
```

```{r}
preds_test <- rep(0, 399)
preds_test[probs_test > 0.5] <- 1
```

```{r}
tb <- table(prediction = preds_test, actual = NBA_test$TARGET_5Yrs)
addmargins(tb)
```

```{r}
roc_obj <- roc(NBA_test$TARGET_5Yrs, probs_test)
```

```{r}
plot(1 - roc_obj$specificities, roc_obj$sensitivities, type="l", 
     xlab = "1 - Specificity", ylab = "Sensitivity")

abline(0, 1, lty=2) # 1-1 line
```

```{r}
auc(roc_obj)
```

```{r}
# comparing two models ROC curve 

roc_model1 <- roc(NBA$TARGET_5Yrs, predict(model1, type = "response"))
roc_model2 <- roc(NBA$TARGET_5Yrs, predict(model2, type = "response"))

# Plot ROC curves
plot(roc_model1, col = "green", main = "ROC Curves", xlim = c(0, 1), ylim = c(0, 1),
     xlab = "False Positive Rate", ylab = "True Positive Rate")
lines(roc_model2, col = "red")
legend("bottomright", legend = c("Model1", "Model2"),
       col = c("green", "red"), lty = 1, cex = 0.8)
```
